### 2014-06-17 paper/ms.tex
rewriting the paper

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
\usepackage{graphics}
\usepackage[dvips]{epsfig}
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman $\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

% definition to produce a "greater than or similar to" symbol
\def\gsim{~\rlap{$>$}{\lower 1.0ex\hbox{$\sim$}}}

\begin{document}

\title{The impact of gas bulk rotation on the Lyman $\alpha$ line} 
\shorttitle{Rotation effects on the Lyman $\alpha$ line}

\shortauthors{Garavito-Camargo, Forero-Romero \& Dijkstra}

\author{ Juan N. Garavito-Camargo, Jaime E. Forero-Romero}
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\email{jn.garavito57@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}
\and
\author{Mark Dijkstra}
\affil{Institute of Theoretical Astrophysics, University of Oslo,
  Postboks 1029, 0858 Oslo, Norway} 
\email{mark.dijkstra@astro.uio.no }

\keywords{galaxies: high-redshift --- line: formation --- methods:
  numerical ---  radiative transfer} 
\begin{abstract}

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$ 
emission line in distant galaxies. 
We model a galaxy as a sphere with an homogeneous mixture of dust and
hydrogen at a constant temperature. 
These spheres have a solid-body rotation with maximum velocities in
the range $0-300$ \kms and neutral hydrogen optical depths in the
range $\tau_{\rm H}=10^{5}-10^{7}$. 
We consider two types of source distributions: single central and a
homogeneoulsy distributed in the sphere. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. 
Observations with a line of sight parallel to the rotation axis yield
line morphologies similar to the static case. 
For lines of sight perpendicular to the rotation axis the intensity at the line
center and the line width increase with rotational velocity.   
Along the same sight, the line becomes single peaked  at rotational
velocities close to half the line width in the static case. 
Notably, we find that rotation does not induce any significant spatial
anisotropy in the integrated line flux nor in the escape fraction. 
We provide an analytical expression for the outgoing \lya spectrum of
a rotating cloud which accounts for the velocity and viewing angle
trends observed in the simulations.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The detection of strong \ly emission lines has become an essential
method in extra-galactic astronomy to find distant star-forming
galaxies
\citep{PartridgePeebles,Rhoads00,Gawiser2007,Koehler2007,Ouchi08,Yamada2012,Schenker2012,Finkelstein2013}.
The galaxies detected using this method receive the 
name of \ly emitters (LAEs). 
A detailed examination of this galaxy population has diverse
implications for galaxy formation, reionization and the large scale
structure of the Universe.  
Attempts to fully exploit the physical information included in the \ly
line require an understanding of all the physical factors involved in
shaping the  line. 
Due to the resonant nature of this line, these physical factors
notably include temperature, density and bulk velocity field of the
neutral Hydrogen in the emitting galaxy and its surroundings.

A basic understanding of the quantitative behavior of the \ly line
has been reached through analytic studies in the case of a static
configurations, such as uniform slabs
\citep{Adams72,Harrington73,Neufeld90} and uniform spheres
\citep{Dijkstra06}. 
Analytic studies of configurations including some kind of bulk flow
only include the case of a sphere with a Hubble like expansion flow
\citep{LoebRybicki}.

A more detailed quantitative description of the \ly line has been
reached through Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. 
In the last two decades these studies have become popular due to the
availability of computing power. 
Early into the 21st century, the first
studies focused on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}. 
Later on, the effects of clumpy media \citep{Hansen06} and
expanding/contracting shell/spherical geometries started to be studied
\citep{Ahn03,Verhamme06,Dijkstra06}.  
Similar codes have applied these results to semi-analytic models of
galaxy formation \citep{Orsi12, Garel2012} and results of large
hydrodynamic simulations \citep{CLARA,Forero12,Behrens13}. 
Recently, Monte Carlo codes have also been applied to the results of
high resolution hydrodynamic simulations of individual galaxies
\citep{Laursen09,Barnes11,Verhamme12,Yajima12}. 
Meanwhile, recent  developments have been focused on the systematic
study of clumpy outflows \citep{DijkstraKramer} and anisotropic
velocity configurations \citep{Zheng2013}.

The recent studies of galaxies in hydrodynamic simulations
\citep{Laursen09,Barnes11,Verhamme12,Yajima12} have all shown
systematic variations in the \ly line with the viewing angle. These
variations are a complex superposition of anisotropic density
configurations (i.e. edge-on vs. face-on view of a galaxy), the
inflows observed by gas cooling and the outflows included in the
supernova feedback process of the simulation. These bulk flows
physically correspond to the circumgalactic and intergalactic medium
(CGM and IGM). These effects are starting to be studied
 in simplified configurations that vary the density and wind
 characteristics \citep{Zheng2013,Behrens2014}.

However, in all these efforts the effect of rotation,
which is an ubiquitous feature in galaxies, has not been
systematically studied. The processing of the \ly photons in a
rotating interstellar medium (ISM) must have some kind of impact in
the \ly line morphology.

Performing that study is the main goal of this paper. We investigate for the
first time the impact of rotation on the morphology of the \ly
line. We focus on a simplified system: a spherical gas cloud with
homogeneous density and solid body rotation, to study the line
morphology and the escape fraction in the presence of dust. We base
our work on two independent Monte Carlo based radiative transfer codes
presented in \cite{CLARA} and \cite{DijkstraKramer}.

This paper is structured as follows: In \S \ref{sec:implementation} we
present the implementation of bulk rotation into the Monte Carlo
codes, paying special attention to coordinate definitions. We also
present a short review of how the \ly radiative transfer codes work
and list the different physical parameters in the simulated grid of
models. In \S \ref{sec:results} we present the results of the
simulations, with special detail to quantities that show a
clear evolution as a function of the sphere rotational velocity. In \S
\ref{sec:discussion} we discuss the implications of our results. In
the last section we present our conclusions. The Appendix presents the
derivation of an analytic expression that can be helpful to interpret
to first order the main trends observed in the Monte Carlo simulations.

In this paper we express a photon's frequency in terms of the
dimensionless variable $x\equiv (\nu -\nu_a)/\Delta\nu_\alpha$, where
$\nu_{\alpha}=2.46\times 10^{15}$ Hz is the Ly$\alpha$ resonance
frequency,  $\Delta\nu_{\alpha} \equiv
\nu_{\alpha}\sqrt{2kT/m_pc^2}\equiv \nu_av_{\rm th}/c $ is the Doppler
broadening of the line which depends on the neutral gas temperature
$T$ scattering the radiation or equivalently the thermal velocity
$v_{\rm th}$ of the atoms.  For the temperature $T=10^4$K used in our
radiative transfer calculations the thermal velocity is
$v_{\rm th}=12.8$\kms.

\section{Models of bulk gas rotation}
\label{sec:implementation}

Describing the kinematics of gas rotation in all generality is a
complex task, specially at high redshift where there is still missing
a thorough observational account of rotation in galaxies beyond
$z>1.0$. Even at low redshifts there is a great
variation in the shape of the rotation curve as observed in HI
emission as a function of the distance to the galaxy center. However
there are two recurrent features. First, in the
central galactic region the velocity increases proportional to the radius,
following a solid rotation behavior. Second, beyond a certain radius
the rotation curve tends to flatten.  An ab-initio description of
such realistic rotation curves in simulations depends on having access to
the dynamic evolution of all mass components in the galaxy: stars, gas
and dark matter. Such level of realism is extremely complex to
achieve, specially if one wants to get a systematic description based
on statistics of simulated objects.

Following the tradition of studies of \ly emitting systems,
we implement a model with simplified geometry. We assume that the gas
is homogeneously distributed in a sphere that rotates as a solid body
with constant angular velocity. This simple model will contain only
one free parameter: the linear velocity at the sphere's surface, $V_{\rm
  max}$.

\subsection{Detailed Implementation of Rotation}

 In the Monte Carlo code we define a Cartesian coordinate system to
 describe the position of each photon. The origin of this system
 coincides with the center of the sphere and the rotation axis is defined
 to be $z$-axis. With this choice, the components of the gas bulk velocity
 field, $\vec{v} = v_{x}\hat{i} + v_{y}\hat{j} + v_{z}\hat{k}$, can be
 written as

\begin{equation}
    v_{x}=-\frac{y}{R}V_{\rm max}, \label{subeq1}
\end{equation}
\begin{equation}
    v_{y}=\frac{x}{R}V_{\rm max}, \label{subeq2}
\end{equation}
\begin{equation}
    v_{z}=0, \label{subeq3}
\end{equation}
%
where $R$ is the radius of the sphere and $V_{\rm max}$ is the linear
velocity at the sphere's surface. The minus/plus sign in the
$x$/$y$-component of the velocity indicates the direction of
rotation. In this case we take the angular velocity in the same
direction as the $\hat{k}$ unit vector. With these definitions we can
write the norm of the angular velocity as $\omega=V_{\rm max}/R$.

For each photon in the simulation we have its initial position inside
the sphere, direction of propagation $\hat{k}_{\rm in}$ and reduced
frequency $x_{\rm in}$. 
The photon's propagation stops once they cross the
surface of the sphere. At this point we store the position, the outgoing direction
of propagation $\hat{k}_{\rm out}$ and the reduced frequency $x_{\rm
  out}$. We now define the angle $\theta$ by $\cos\theta = \hat{k}_{\rm out}\cdot
\hat{k}\equiv \mu$, it is the angle of the outgoing photons with 
respect to the direction of the angular velocity. We use the variable $\mu$ to
study the anisotropy induced by rotation. Figure \ref{fig:geometry}
shows the geometry of the problem and the important variables.

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f1.pdf}
\end{center}
\caption{Geometry of the gas distribution. The angular velocity vector
  is parallel to the unit vector $\hat{k}$. In order to describe the
  departures from spherical symmetry we use the polar angle $\theta$
  formed by the direction of the outgoing photons with respecto to the
  $z$-axis. We define define the variable $\mu\equiv\cos\theta$ to
  report to present our results. Computing the spectra for photons in
  a   narrow range of $\mu$ is equivalent to having a line-of-sight
  oriented in that direction.  
    \label{fig:geometry}}  
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f2.pdf}
\end{center}
\caption{
2D histogram showing the number of photons that escape with frecuency
$x$ forming an angle $\theta$ (parameterized as $|\cos\theta|$) with the
rotation axis. 
The rotational velocity ($0,100,200,300$\kms) increases from left to
right and the optical depth ($10^5$, $10^6$, $10^7$) from top to
bottom. 
The \ly photons are initialized at the center of the sphere. 
Two main results can be read from this figure.
First, the line morphology depends on the viewing angle. 
Second, the line can become single peaked for high rotational
velocities.
\label{fig:CentralSpec} }   
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f3.pdf}
\end{center}
\caption{Same as Figure \ref{fig:CentralSpec} for \ly photons
  initialized homogeneously throughout the sphere.
    \label{fig:HomSpec}}  
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f4.pdf}
\end{center}
\caption{Shape of the \ly line for different maximum rotational
  velocities for a LoS perpendicular to the rotation axis
  ($|\mu|\sim 0$). The continuous (dashed) line represents the central
  (homogeneous) source distributions.  The panels follow the same
  distribution as in Figures \ref{fig:CentralSpec} and \ref{fig:HomSpec}.
    \label{fig:differentvelocities}}  
\end{figure*}

\subsection{Brief Description of the Radiative Transfer Codes}

Here we briefly describe the relevant characteristics of the two
radiative transfer codes we have used.  
For a detailed description we refer the reader to the original papers
\cite{CLARA,DijkstraKramer}.

The codes follow the individual scatterings of \ly photons as they
travel through a 3D distribution of neutral Hydrogen. 
The frequency of the photon (in the laboratory frame) and
its direction of propagation change at every scattering. 
This change in frequency is due to the peculiar velocities of the
Hydrogen absorbing and re-emitting the photon. 
Once the photons escape the gas distribution we store their direction of
propagation and frequency at their last scattering.

The initialization process for the \ly photons especifies its
position, frequency and direction of propagation. 
We select the initial frequency to be exactly the \ly rest-frame
frequency in the gas reference frame and the direction of propagation
to be random following an flat probability distribution over the sphere.  
It means that for photons emitted from the center of the sphere
$x_{\rm in}=0$, while photons emitted at some radii with a peculiar
velocity $\vec{v}$ have initial values $x_{in}$ depending on its direction of
propagation: $x_{\rm   in}=\vec{v}\cdot\hat{k}_{\rm in}/v_{\rm th}$.
We do not include the effect of turbulent velocities in the
initialization.
We neglect this given that the induced perturbation should be on the
close to the thermal velocity, $12.8$\kms, which is one order of
magnitude smaller than the velocity widths ($100$-$500$\kms) in the
static case.

If dust is present, the photon can interact either with a Hydrogen
atom or dust grain.
In the case of a dust interaction the photon can be either absorbed or
scattered.
This probability is encoded in the dust albedo, $A$, which we chose
to be $1/2$. 
In order to obtain accurate values for the escape fraction of
photons in the presence of dust, we do not use any accelerating
mechanism in the radiative transfer.

The codes treat the gas as homogeneous in density and temperature.  
This implies that the cas is completely defined by its geometry
(i.e. sphere or slab), temperature $T$, Hydrogen optical depth
$\tau_{\rm H}$, dust optical depth $\tau_{\rm a}$ and the bulk
velocity field $\vec{v}$.

\subsection{Grid of Simulated Galaxies}
\label{sec:models}

In the Monte Carlo calculations we follow the propagation of $N_{\gamma}=10^5$
numerical photons through different spherical galaxies. 
For each galaxy we vary at least one of the following parameters: the maximum
rotational velocity $V_{\rm max}$, the hydrogen optical depth $\tau_{H}$,
the dust optical depth $\tau_{a}$ and the initial distribution of photons
with respect to the gas. 
In total there are $48$ different models combining all the possible
different variations in the input parameters.
Table \ref{table:models} lists the different parameters we used to
generate the models.

Additionally, we have used two independently developed Monte Carlo
codes \citep{CLARA,DijkstraKramer} to perform the calculations. 
The results and trends we report are observed in both codes.

\begin{table}
\begin{center}
\begin{tabular}{llc}\hline\hline
Physical Parameter (units) & Symbol & Values\\\hline
Velocity (\kms) & $V_{\rm max}$&$0,\ \ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{a}$ & $0$,$1$\\
Photons Distributions & & Central, Homogeneous\\\hline\hline
\end{tabular}
\caption{
  Summary of Physical Parameters of our Monte Carlo Simulations.} 
\label{table:models}
\end{center}
\end{table}

\section{Results}
\label{sec:results}

The central result of this paper are summarized in Figures
\ref{fig:CentralSpec} and \ref{fig:HomSpec}. 
They show 2D histograms of the escape frequency $x$ and outgoing angle
$\theta$ parameterized by $|\mu|$. 
Taking into account only photons photons around a value
of $|\mu|$ gives us the emission detected by an observer located at an
angle $\theta$ with respect to the rotation axis. 
We have veryfied that the solutions are indeed symmetric with respect
to $\mu=0$.

From these Figures we can see that the line properties change with
rotational velocity and depend on the viewing angle $\theta$.  
In the next subsections we quantify the morphology changes with with
velocity, optical depth and viewing angle.  
We characterize the line morphology by its total intensity, the full
width at half maximu, (FWHM) and the location of the peak maxima. 
In order to interpret the
morphological changes in the line we also report the median number of
scatter for each \ly photon in the simulation. 
For the models where dust is included we measure the escape fraction
as a function of rotational velocity.

\subsection{Line Morphology}
\label{sec:angles}

The first column in both Figures \ref{fig:CentralSpec} and
\ref{fig:HomSpec} shows that for the static sphere the line properties
are independent of $|\mu|$, as it is expected due
to the spherical symmetry. 
However, for increasing rotational velocities, at a fixed optical
depth, there are clear signs that this symmetry is broken.

If the viewing angle is aligned witht the rotation axis, $|\mu|\sim
1$, the \ly line keeps in most of the cases a double peak with minor
changes as the rotational velocity starts to increase. 
However, for a line of sight perpendicular to the rotation axis,
$|\mu|\sim 0$, the impact of  rotation is larger.  
The double peak readily transforms into a single peak.

This is clear in  Figure \ref{fig:differentvelocities} where we
present the different line morphologies for $|\mu|\sim 0$ in the
homogeneous and central configurations.  
The panels have the same distribution as Figures \ref{fig:CentralSpec}
and \ref{fig:HomSpec}.    
There are three clear effects on the line morphology as the rotational
velocity increases.  
First, the line broadens; second, the double peaks reduce their intensity; and
third, the intensity at the line centre rises. 
The last two effects are combined to give the impression that the double
peaks are merged into a single one at high rotational velocities.

\subsection{Integrated Line Intensity}

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f5.pdf}
\end{center}
\caption{Integrated flux distribution as a function of the
  viewing angle as parameterized by $\mu$. Continous (dashed)
  correspond to central (homogeneous) source distribution.
  The models correspond to an optical depth of $\tau_{\rm
    H}=10^5$ and rotational velocities of $100$\kms, $200$\kms and
  $300$\kms. The distributions are flat in the range of models probed
  in this paper, meaning that the integrated flux for all viewing
  angles is the same.
\label{fig:muhisto}} 
\end{figure}

We now consider possible variations in the integrated flux with
respect to the viewing angle $\theta$. 
To this end we  define the normalized flux seen by an observer at an
angle $\mu$ by:

\begin{equation}
F(\mu) = \frac{2\Delta N}{N\Delta \mu}, 
\end{equation} 
%
where $\mu=\cos\theta$, $N$ is the total number of outgoing photons,
$\Delta N$ is the number of photons in an angular bin $\Delta
\Theta$. This definition satisfies the condition
$\int_{-1}^{1}F(\mu)d\mu/2=1$.  In the case of perfect spherical
simmetry one expects a flat distribution with $F(\mu)=1$.

Figure \ref{fig:muhisto} shows the results for a selection of models
with $\tau_{\rm H}=10^{5}$, ifferent rotational velocities and the two
types of source distributions.
This Figure shows that $F(\mu)$ is consistent with being flat, apart
from some statistical fluctuations on the order of \%2.

This is a remarkable result. 
In spite of that the rotation axis defines preferential direction, the
integrated flux is the same for all viewing angles, at least for the
range of parameters explored in this paper.

\subsection{Full Width at Half Maximum}
\label{sec:widthpeak}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f6.pdf}
\end{center}
  \caption{FWHM for the non-dusty models as a function of the viewing
  angle parameterized by $|\cos\theta|$. Continuous (dashed) lines  correspond
  to central (homogeneous) source distributions. The general trend is
  of an decreasing line width as the line of sight becomes parallel to the
  rotation axis.
  \label{fig:widthvsmu}} 
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f7.pdf}
\end{center}
\caption{FWHM for the non-dusty models as a function of
  rotational velocity $V_{\rm max}$ for observers located
  perpendicular to the rotation axis. 
  The left panel shows the results in velocity units while the right
  panel normalizes the data by the FWHM in the static case. 
  Continuous (dashed) lines  correspond to central (homogeneous)
  source distributions. 
  The straight lines represent the fit to the data using the
  expression in Eq. (\ref{eq:fwhm}). 
  \label{fig:widthsvsvelocity}}
\end{figure*}

We use the full width at half maximum (FWHM) to quantify the line
broadening. 
We measure this width from the line intensity histogram by finding the
values of the velocities at half maximum intensity.
We use lineal interpolation between histogram points to get a value
more precise than the bin size used to construct the histogram.

Figure \ref{fig:widthvsmu} shows the FWHM for all models as a function
of the viewing angle. 
The FWHM increases for decreasing values of $\mu$ (movement from the
poles to the equator) and increasing values of $V_{\rm max}$. 
In Figure
\ref{fig:widthsvsvelocity} we fix $|\mu|<0.1$, i.e. viewing angle
perpendicular to the rotation axis, to plot the FWHM as a function of
rotational velocity.

Ee parametrize the dependency of the line width with  $V_{\rm max}$ as
%
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the velocity width in the static case and $\lambda$ 
is a positive scalar to be determined as a fit to the data.  
With this test we want to know to what extent the new velocity width can be
expressed as a quadratic sum of the two relevant velocities in the
problem.

All the models fall into a single familiy of lines in the plane shown
in the right panel of Figure \ref{fig:widthsvsvelocity}, justifying
the choice of our parameterization.
We fit simultaneously all the points in two separate groups, central
and homogeneous sources.
We find that these values are $\lambda_{\rm c} = 0.83 \pm 0.06$ and
$\lambda_{\rm    h}= 0.82\pm 0.05$ respectively.

\subsection{Line Maxima}
\label{sec:maxima}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f8.pdf}
\end{center}
\caption{Position of the line maxima as a function of maximum
  rotational   velocity $V_{\rm max}$. Continuous (dashed) lines
  correspond to   central (homogeneous) source distributions. A value
  of $x_{\rm     max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}
\end{figure*}

We measure the peak maxima position to quantify the transition from
double into single peak profiles. We do this as a function of the viewing angle
$\theta$ for different rotational velocities. The results are
summarized in Figure  \ref{fig:maximumsvsvelocity}. There are two
interesting features that deserve attention. First, for a viewing angle
parallel to the rotational axis ($\mu\sim 1.0$) the maxima of all
models are similar regardless of the rotational velocity. Second, at a
viewing angle perpendicular to the rotation axis ($\mu\sim 0.0$) a
good fraction of models become single peaked. This feature appears
more frequently for homogeneously distributed sources if all the
other parameters are equal.

\subsection{Dusty Clouds: Escape Fraction}
\label{sec:escapefraction}

\begin{table}
\begin{center}
\begin{tabular}{c ccc ccc}
\hline \hline
Source & $\tau_{H}$ & & & $\ V_{\rm max}$& & \\
Distribution& &  &  & (\kms) & & \\ 
& & 0 & 50 & 100 &200 & 300\\ \hline 
Homogeneous & $10^{5}$& 0.263 & 0.266 &  0.309 &  0.357 &  0.370  \\
            & $10^{6}$ & 0.289 &  0.289 &   0.309 &  0.360 &  0.393 \\
            &$10^{7}$ &  0.227 &  0.229 &  0.231 &  0.254 &  0.281 \\
Central & $10^{5}$ &  0.096 & 0.097 & 0.096 &  0.096 & 0.096 \\
  		&$10^{6}$ & 0.066 & 0.066 &  0.066 &  0.066 &  0.066 \\
 		&$10^{7}$ & 0.015 &  0.016 & 0.016 & 0.016 & 0.015 \\
\hline
\end{tabular}
\caption{
 Escape fraction values for all dusty models. } 
\label{table:escape}
\end{center}
\end{table}

%\begin{figure*}
%\begin{center}
%  \includegraphics[width=0.40\textwidth]{f10.pdf}
%  \includegraphics[width=0.40\textwidth]{f11.pdf}
%\end{center}
%  \caption{Effects of rotation on the escape fraction. Left
%    panel, escape fraction as a function of rotational velocity. All
%    these models have $\tau_{a}=1$. The continuous (dashed) lines
%    correspond to central (homogeneous) models. Right panel, escape
%    fraction as a    function of the product $(a\tau_{\rm
%      H})^{1/3}\tau_{A}$. The    analytic solution for the infinite
%    slab is shown as a continuous line. Different lines correspond
%    to different rotational velocities for the homogeneous models.     
%    \label{fig:escape_fraction}} 
%\end{figure*}

We now estimate the escape fraction $f_{\rm esc}$ for the dusty
models. The main  result is that we do not find any dependence
with the viewing angle but there is a strong dependence with the
rotational velocity for the homogeneous models.

The left panel in Figure \ref{fig:escape_fraction} summarizes the
findings about the dependence with the rotational velocity. The left
panel shows the escape fraction as a function of the rotational
velocity. To construct this Figure we take into account all the
photons regardless of its outgoing direction. We observe that the 
curves for the central source distribution stay flat, while for the
homogeneous case there is a clear rise with rotational velocity.
Rotation has a higher relative impact in the models with low optical
depth. For instance, in models with $\tau_{\rm H}=10^5$, the static
escape fraction is $0.26$ and increases to $0.37$ for $V_{\rm
  max}=300$ \kms. Table \ref{table:escape} lists all the values for
the escape fraction.

In the right panel of Figure \ref{fig:escape_fraction} we put these
results in the context of the analytic solution for the infinite slab
\citep{Neufeld90}. In Neufeld's set-up the analytic solution depends
uniquely on the product $(a\tau_{\rm   H})^{1/3}\tau_{A}$ where
$\tau_{A} = (1 - A)\tau_{a}$, valid only in the limit $a\tau_{\rm
  H}\gg 1$. The dashed lines in  the right panel of Figure
\ref{fig:escape_fraction} show the results for the different
rotational velocities for the homogeneous models. First of all we note
that the escape fraction increases slightly from $\tau_{\rm H}=10^5$ to
$\tau_{\rm H}=10^{6}$, which is counterintuitive as we expected that increasing
$\tau_{\rm H}$ requires more scattering before the photon escapes, which would
increase the probability that photons are destroyed by dust.

However, this naive interpretation is incorrect: increasing $\tau_{\rm
  H}$ from $10^{5}$ to $10^{6}$ causes us to transition from the
'optically thick' to  the 'extremely optically thick'
  regime. In the optically thick regime, Lya  photons can escape in
  'single flight' which corresponds to a scenario in which the
  photon resonantly scatters $10^4-10^5$ times until it is scattered
  into the wing of the line ($x\sim 3-4$). At these frequencies the
  medium is optically thin, and the photons can escape efficiently in
  a single flight. In contrast, in an extremely optically thick medium \ly
  photons escape in a `single excursion' \citep{Adams72}. Here, photons
that are scattered into the wing of the line escape from the medium in
a sequence of wing scattering events.In both cases, \ly photons
resonantly scatter $10^4$-$10^5$ times. Because we keep our clouds the same
size, the mean free path of Lya photons that scatter resonantly is 10
times larger for the case $\tau_{\rm H}=10^5$ than for $\tau_{\rm
  H}=10^6$. If we compute the average distance $D$ travelled by \ly
photons through a medium of size $R$ as a function of line center
optical depth $\tau_{\rm H}$, then we find that during the transition
from optically thick to extremely optically thick the mean traversed
distance $D$ actually decreases slightly.  This decrease is unique to
this transition region, and $D$ generally increases with $\tau_{\rm H}$ at other
values of $\tau_{\rm H}$.

\subsection{Average Number of Scatterings}
\label{sec:scatterings}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{f14_1.pdf}
  \includegraphics[width=0.48\textwidth]{f14_2.pdf}
  \includegraphics[width=0.48\textwidth]{f14_3.pdf}
  \includegraphics[width=0.48\textwidth]{f14_4.pdf}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left panels 
      corresponds to the static case and right panels correspond to
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}. }
\end{figure*}

The number of scatterings affects the escape frequency of a \ly
photon. Studying this quantity can help us in claryfying two results
that we have observed for homogeneous models in the previous sections:
the quick emergence of a single peak and the increase in the escape
fraction.

A possible explanation for these two results is that some photons
close to the surface can escape with a low number of scatterings. This
has a two consequences: the escaping photons do not have time to
scatter away from the line's center and also have less chances to find
an absorbing dust grain.

In Figure \ref{fig:Nscatt} we show the average number of scatterings
$\langle N_{\rm scatt}\rangle$ as a function of the cosinus of the
outgoing angle $|\cos\theta|$ and the rotational velocity
$V_{\rm max}$. From the right panel observe that the number of
scatterings and the outgoing angle are independent. This plot
corresponds to the specific case of the central model with $\tau=10^5$ and
$V_{\rm max}=300$\kms, but we have veryfied that this holds for all
models. This allows us to construct statistics using all the outgoing
photons regardless of their value of $\cos\theta$.

The right panel of Figure \ref{fig:Nscatt} shows how the
average number of scatterings decreases for larger rotational
velocities in the case of homogeneous source distributions. For instance, for
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$61\%$ at $V_{\rm max}=300\kms$ in comparison to the static case. For
the central source distribution the average number of scatterings
$\langle N_{\rm   scatt}\rangle$ changes less than $0.5\%$.

In order to gain a deeper understanding of these results  we make 2D
histograms for the number of scatterings as a function of the outgoing
dimensionless frequency $x$. In Figure \ref{fig:Nscatt2D} we show these 
results in the case $\tau_{\rm H}=10^5$ for the
static case and $V_{\rm max}=300$\kms. The upper (lower) panels show the
results for the homogeneous (central) source distribution. The color
scale is logarithmic in the number of photons at a certain value
$x-N_{\rm scatt}$.

The top right panel of Figure \ref{fig:Nscatt2D} (homogeneous sources,
high rotational velocity) supports our hypothesis about
the photo-sphere in the homogeneous distribution. In this case most of
the photons that left with $x\sim 0$ have escaped with less than $10$
scatterings. This also explains the decrease in the average number of
scatterings observed in Figure \ref{fig:Nscatt}. We have also veryfied
that these photons are indeed emitted closer to the surface than to
the center of the sphere.

However, for a central distribution the situation is quite different
(lower panels in Figure \ref{fig:Nscatt2D}). In this case the number
of scatterings remains high, in the order of the optical depth, but
the two peaks do get closer to each other. The most probable
physical picture is that each scattering, due to the bulk velocity of
the gas, is inefficient in driving the photon outside the line
center.

We now proceed with a more detailed quantification of the number of
scatterings in our simulations. We guide the measurement by the fact
that in static cases, a large value of the optical depth correlates
with a high number of scatterings.  This has been precisely quantified
in the case of static infinite slab. In this model for centrally
emitted sources the average number of scatterings depends only on the
optical depth $\langle N_{\rm  scatt}\rangle=1.612\tau_{\rm   H}$
\citep{Adams72,Harrington73}, for homogeneously distributed sources
$\langle N_{\rm   scatt}\rangle=1.16\tau_{\rm   H}$
\citep{Harrington73}.

We perform a similar quantification of in our speherical setup for the
static spheres. We find that for the central model the number of
scatterings is proportional to the optical depth, with $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$ for optical depth
values of $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$ respectively.
For the homogeneous static sphere we find that $\langle N_{\rm
  scatt}\rangle= (0.99, 0.59, 0.51)\tau_{\rm   H}$, this represents
factors of $(0.66, 0.59, 0.55)$ lower than the centrally  emitted photons.

For the homogeneous rotating sphere at  $V_{\rm{max}}=300\kms$ we find
that $\langle N_{\rm   scatt}\rangle=(0.38, 0.34, 0.42)\tau_{\rm   H}$
which represents   factors of $(0.26, 0.34, 0.46)$ lower than the
corresponding central case.

\section{Discussion}
\label{sec:discussion}

We now compare our findings to other computational results and discuss
its possible implications  for the interpretation of observational data.

{\bf Galaxy simulations with gas rotation}. \cite{Verhamme12} studied \ly
line emission in two high resolution simulations of individual
galaxies. The main purpose of their study was to assess the impact of
two different Inter-Stellar Medium (ISM) prescriptions. However, each
simulated galaxy had a disc structure with a clear rotation pattern in
the ISM and inflowing gas from the circum-galactic region. The configuration
had an axial symmetry and they reported a strong dependence of both
the escape fraction and the total line intensity as a function of the
$\theta$ angle. In our results, none of these two quantities has a
dependence either on the inclination angle or the rotational
velocity. We suggest that he effect reported by \cite{Verhamme12} is
consistent with being a consequence of the different hydrogen optical
depth for different viewing angles and not as an effect of the bulk
rotation.

{\bf Single peaked lines}. The presence of single peaked profiles has
been associated to inflow/outflow dynamics
\citep{Verhamme06,DijkstraKramer}. In this paper we show that gas bulk
rotation can also be considered as a probable origin for that
behavior, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be
considered in the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. This highlights
that in order to interpret the observations of \ly emitting galaxies
it is necessary to consider the possible effect of different
inclination angles along the line of sight.

{\bf Systemic velocities}. There are observational measurements for the
velocity shift between the \ly and other emission lines. In our study
we find that the position of the peak maxima is insensitive to the
rotational velocity of the gas. In our model the only parameter that
can influence the peak maxima position is the optical depth.

{\bf Anisotropic emission}. In our studies the integrated flux is
constant is independent of the viewing angle. This rules out the
possibility of having an observational indication of gas bulk rotation
from measurements of the line equivalent width. A possible way to
obtain an anisotropic flux with the same gas configuration is to allow
for anisotropic emission \citep{Zheng2013,Behrens2014}. We explore
these scenarios in future work (Remolina-Gutierrez et al., in prep.).

{\bf Additional uncertainty on the \ly escape fraction}. Study of
high redshift LAEs in numerical simulation often requires the
estimation of the \ly escape fraction in order to compare their
results against observations
\citep{CLARA,Dayal2012,Forero12,Orsi12,Garel2012}. Most of these
models estimate the escape fraction from the column density of dust and
neutral Hydrogen. The results of our simulation suggest that the
rotational velocity has also to be accounted for, having an impact on
the values of the escape fractions by factors of $\sim 2$.

\section{Conclusions}
\label{sec:conclusions}

In this paper we quantified for the first time in the literature the effects
of gas bulk rotation in the morphology of the \ly emission line in star forming
galaxies.  Our results are based on the study of an homogeneous sphere
of gas with solid body rotation. We explore a range of models by varying
the rotational speed, hydrogen optical depth, dust optical depth and
initial distribution of \ly photons with respect to the gas
density. As a cross-validation, we obtained our results from two
independently developed Monte-Carlo radiative transfer codes.

Two conclusions stand out from our study. First, rotation clearly
impacts the \ly line morphology; the width and the relative intensity
of the center of the line and its peaks are affected. Second, rotation introduces an
anisotropy for different viewing angles. For viewing angles close to
the poles the line is double peaked and it makes a transition to a
single peaked line for high rotational velocities and viewing angles
along the equator. This trend is clearer for spheres with
homogeneously distributed radiation sources than it is for central
sources.

Remarkably, we find two quantities that are invariant with respect
to the viewing angle and the rotational velocity: the integrated flux and the escape
fraction.  The exception is the models with homogeneously distributed
sources, which present an escape fraction dependent on the rotational
velocity, but still independent of the viewing angle. We showed that
these results could be interpreted in terms of the average number of
scatterings for the photons.

Quantitatively, the main results of our study are summarized as
follows.

\begin{itemize}

\item In all of our models, rotation induces changes in the line morphology
 for different values of the angle between the rotation
 axis and the LoS, $\theta$.  The changes are such that for 
 $\cos\theta\equiv \mu\sim 0$, i.e. LoS perpendicular to the
 rotation axis, and high rotational velocities the line becomes single peaked.

\item For a viewing angle with $\mu=0$ the line width increases with rotational
  velocity. This change approximately follows the functional form  ${\rm FWHM}^2 = {\rm FWHM}_{ 0}^2 + (V_{\rm max}/\lambda)^2$, where FWHM$_{0}$ indicates the line
  width for the static case and $\lambda$ is a constant. We have
  determined this constant to be  $\lambda_{\rm c}=0.84 \pm 0.06$ and
  $\lambda_{\rm h}=0.54\pm 0.10$ for the central and homogeneous source
  distributions, respectively.

\item At fixed rotational velocity the line width decreases as $|\mu|$
  increases, i.e. the smallest value of the line width is observed for
  a line of sight parallel to the ration axis. However the lowest
  values are always larger than the values for the static case.

\item The single peaked line emerges for homogeneously distributed sources 
  when the rotational velocity is larger than half FWHM$_0$ and $\mu\sim
  1$.

\item  For central sources we find that the escape fraction does not change
  with rotational velocity nor with the viewing angle. We interpret this
  in terms of the average number of scatterings which does not
  significantly change with rotation. This is due to the  
  fact that the vast majority of scatterings events are resonant, 
  during which the mean free path is very short and the effect of gas
  bulk rotation is not enough to affect \ly photons.

\item For homogeneous sources we find that the  as rotational velocity
  increases, the escape fraction rises by a factor of $\sim 2$ with
  respect to the static situation. We also interpret this change in
  terms of the average number of scatterings. In the second part of \S
  \ref{sec:widthpeak} we show that the average number of scatterings
  decreases a $40\%$ compared to the static case, while a large
  fraction of photons escape with less than $10$ scatterings.

\end{itemize}

Comparing our results with recent observed LAEs we find that many 
morphological features such as high central line flux, single peak
profiles  could be explained by gas bulk  rotation present in these
LAEs.

The definitive and clear impact of rotation on the \ly morphology
suggests that this is an effect that should be taken into accunt at
the moment of interpreting high resolution spectroscopic data. To this
end, we defer a discussion of observed lines features in the context of gas
bulk rotation to a future publication (Remolina-Gutierrez et al., in
prep.).

\section*{Acknowledgments}

JNGC acknowledges financial support from Universidad de los
Andes.

JEFR acknowledges financial support from Vicerrectoria de
Investigaciones at Universidad de los Andes through a FAPA grant.

We thank the International Summer School on AstroComputing
2012 organized by the University of California High-Performance
AstroComputing Center (UC-HiPACC) for providing computational
resources where some of the calculations were done.

The data, source code and instructions to
replicate the results of this paper can be found
here {\texttt{https://github.com/jngaravitoc/RotationLyAlpha}}.
Most of our code benefits from the work of the IPython and Matplotlib
communities \citep{IPython,matplotlib}.

We thank the referee for the suggestions that allowed us to greatly
improve and better frame the interpretation of our simulations.

\appendix
\section{Analytic Expression for the Ly$\alpha$ Spectrum
  emerging from Rotating Cloud}

Ly$\alpha$ scattering through a an optically thick gas cloud that is
undergoing solid-body rotation (i.e. the angular speed around the
rotation axis is identical for each hydrogen atom) proceeds identical
as in a static cloud. 
While scattering events off atoms within the rotating cloud impart
Doppler boosts on the Ly$\alpha$ photon, these Doppler boost are only
there in the lab-frame. 
In the frame of the rotating gas cloud, all atoms are stationary w.r.t
each other, an the scattering process proceeds identical as in the static
case\footnote{An alternative - more quantitative way - to view this is
  to consider the path of an individual photon: say the photon travels
  a distance corresponding to $\tau=1$, scatters by $90$ degrees, then
  travels another distance corresponding to $\tau=1$. It is easy to
  show that the point where the photon ends up is independent of the
  assumed rotation velocity of the cloud.}. 
This explains why the escape fraction of Ly$\alpha$ photon is
independent of the rotation velocity off the cloud.

The spectrum of \lya photons emerging from a rotating gas cloud is
also identical as for the static case in a frame that is co-rotating
with the cloud. However, the surface of cloud now moves in the
lab-frame. Each surface-element on the rotating cloud now has a bulk
velocity w.r.t a distant observer. In order to compute the spectrum
emerging from this rotation cloud, we have to sum over all surface
elements (weighted by their intensity).

To provide an analytic description of this emerging spectrum, we adopt
the geometry shown in Fig~\ref{fig:scheme}. 
Note that this differs from the scheme in Fig~1 from the main body of
the paper.

\begin{figure*}[h]
\centerline{\includegraphics[width=85mm]{scheme1.pdf}
\includegraphics[width=85mm]{scheme2.pdf}}
\caption[]{Adopted geometry for evaluating the analytic spectrum.}
\label{fig:scheme}
\end{figure*}

The sightline to the observer \& rotation define the $x-z$ plane. 
The {\it left panel} in Fig~\ref{fig:scheme} shows the view from the $y$-axis. 
The observer sits along the $x$ axis. 
The rotation axis makes an angle $i$ with respect to the $z$-axis. 
We sum up spectra from individual patches by integrating over the
impact parameter $b$, and angle $\phi$. 
Each $(b,\phi)$ corresponds to a point on the sphere. 
This point has a velocity vector ${\bf v}(b,\phi,i)$, which we denote
with ${\bf v}$ for brevity.

The intensity at the surface point $(b,\phi)$ is

\begin{displaymath}
J(x,b,\phi,i)=\frac{\sqrt{\pi}}{\sqrt{24}a\tau_0}\Bigg{(}\frac{(x-x_{\rm
    b})^2}{1+{\rm cosh}\Big{[}\sqrt{\frac{2\pi^3}{27}}\frac{|(x
      -x_{\rm b})^3|}{a\tau_0}\Big{]}}\Bigg{)},
\end{displaymath} 
%
where $x_b\equiv v_{b}/v_{\rm th}$, and $v_b$ is the component of ${\bf v}$ projected onto the line-of-sight. This component is given by
\begin{equation}
v_{\rm b}(b,\phi,i)=V_{\rm max}\sqrt{1-(b/R)^2\cos^2 \phi}\hs \cos i
\hs \sin a,
\end{equation}
%
where

\begin{equation}
\tan a = \frac{ b\sin \phi}{\sqrt{R^2 -b^2}}.
\end{equation}

In order to compute to total intensity we integrate over $b$ and
$\phi$ with a weight wigen by the surface brightness of the
sphere at $(b,\phi)$, $S(b,\phi)$.

\begin{displaymath}
J(x,i)=2\pi \int_0^Rdb \hs b \int_0^{2\pi}d\phi \hs
S(b,\phi)J(x,b,\phi,i) \approx 2\pi \int_0^Rdb \hs b
\int_0^{2\pi}d\phi \hs J(x,b,\phi,i)\\ \nonumber.
\end{displaymath}
%
In the last expression we assume that $S(b,\phi)$ is constant. 
This corresponds to $I(\mu) \propto \mu$ at the surface, where $\mu$
denotes the cosine of the angle of the propagation direction of the
outgoing photon and the normal to the spheres surface: a fixed $db$
corresponds to a physical length $ds = db/\mu$ on the sphere.  
If $I(\mu)$ were constant, this would imply that the sphere should
appear brighter per unit $b$. 
A constant surface brightness profile requires the directional
dependence for $I(\mu) \propto \mu$ to correct for this. 
Indeed, this is what is expected for the escape of Ly$\alpha$ photons
from static, extremely opaque media (see \cite{Ahn01}; their
Fig~4 and accompanying discussion).

It is worth stressing that the analytic calculation is clearly not a
complete analytic calculation, and we do not expect perfect agreement:
we {\it assumed} a functional form for the surface brightness profile
[or for $I(\mu)$].

It is important to realise that $I(\mu)$ itself may
depend on the spectrum. I.e., analytic solutions exist for $J(x) =
\int_0^1 I(x,\mu) d\mu$ at the boundary, and {\it approximate}
expressions for $I(\mu) =\int dx I(x,\mu)$, but none for $I(x,\mu)$
itself.

Fig~\ref{fig:comparison}shows some examples of analytic vs. full MC
spectra.
The area under the spectra is the same in all cases. 
The left panel shows the results for different rotational velocitiies
in the case of $\tau_{H}=10^7$ and an observer located perpendicular
to the axis of rotation ($\mu=0$ in the notation of the main body of
the paper, $i=0$ in the scheme of Fig~\ref{fig:scheme}). 
The right panel shows the results for different viewing angles in the
case of $\tau_{H}=10^7$ and a rotational velocity of $V_{\rm
  max}=300$\kms.

The two methods clearly give good agreement, though not perfect. In
particular, the left panel shows that the MC gives rise to a spectrum that is
slightly more concentrated towards the line centre. 
It is easy to get better agreement my modifying the surface brightness
profile (e.g. if $S(b)=2-b$, then the analytic calculation agrees
better).
In any case, the analytic calculation closely captures the results
obtained from the full calculations from the MC simulations.  
As such, they are extremely useful and provide us with a quick tool to
verify our calculations at the first order level.

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{vary_vel.pdf}
  \includegraphics[width=0.48\textwidth]{vary_angle.pdf}
\end{center}
\caption{
Comparison of the Monte Carlo results agains the analytic
solution. The left panel explores the results of different velocities.
The right panel presents the results for two different observers:
paralel and perpendicular to the rotationa axsi, $\mu=1$ and $\mu=0$
respectively.
\label{fig:comparison} }   
\end{figure*}

\bibliographystyle{apj}
\bibliography{references}

\end{document}