### 2014-09-08 paper/referee-reply.tex
new referee reply

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{graphicx}

\title{Response to the second report on the article draft \emph{The
    Impact of Gas Bulk Rotation on the Lyman-$\alpha$ line.}} 
\author{Juan N. Garavito-Camargo, Jaime E. Forero-Romero, Mark Disjkstra}
\date{\today}

\begin{document}

\maketitle

We thanks the referee for the new minor comments. 
In what follows the comments by the referee are boldfaced.

Best regards, \\

The Authors\\

\section*{Minor Comments}

{\bf About the directionallity of the escaping photons, why did you select only photons with $0 < \varphi < \pi/2$? Did you then normalise the blue and green curves to sample the same number of photons ? Now that I understood that Lya transfer in solid-body rotating cloud proceeds like in a static medium, until projection in an external, non-rotating frame, the only expected trend should be with optical depth, indeed. These curves show $I(\mu)$ that you discuss in the appendix, right ? It seems that the regimes that you investigated are not yet enough optically thick so that $I(\mu) = \propto \mu$. Maybe you can show these plots on the appendix, to illustrate that the analytic solution becomes better in veryoptically thick regimes.}

We did not limit the selection to $0 < \varphi < \pi/2$. The outgoing
photons close to the surface follow are in that range, where $\varphi$
describes the angle with respect to the normal to the surface.  The
curves have the same normalization in each of the plots.

We have included these plots in the appendix and extended the
discussion in the direction suggested by the referee.

\end{document}